---
title: "Accessing Vectorbase in R using biomaRt"
author: "Marion Shadbolt"
date: "26 August 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing biomaRt

If you haven't already installed biomaRt use these commands:

```{r eval=F}
## try http:// if https:// URLs are not supported
source("https://bioconductor.org/biocLite.R")
biocLite("biomaRt")
```

## Viewing available biomart databases

By default, biomaRt thinks you are interested in ensembl genes, it sets your `host = "www.ensembl.org"` so that if you follow the commands in the guide you get:

```{r}
library(biomaRt)
listMarts()
```

Which isn't great because the vectorbase databases are not listed, to see what vectorbase databases are available we need to change the host to `biomart.vectorbase.org` like this:

```{r}
listMarts(host = "biomart.vectorbase.org")
```

You can then choose which database you are interested in, for example let's choose the `"vb_gene_mart_1608"`

```{r}
vectorbase_gene <- useMart("vb_gene_mart_1608", host = "biomart.vectorbase.org")
listDatasets(vectorbase_gene)
```

## Storing a dataset for further use
So we can see all the available genesets for the organisms on hosted on vectorbase. If we want the *Aedes aegypti* dataset:

```{r}
aaegGenes<-useDataset("aaegypti_eg_gene", vectorbase_gene)
aaegGenes
```

If you already know that you're only interested in *Ae. aegypti* genes you can skip a few of the steps above by directly using the following commands:
```{r}
aaegGenes<-useMart(biomart = "vb_gene_mart_1608", dataset = "aaegypti_eg_gene", host = "biomart.vectorbase.org")
aaegGenes
```