---
title: "Mining GO ids for Gowinda from BioMart"
author: "Marion Shadbolt"
date: "4 September 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

To get an up to date version of Gene ontology terms for _Aedes aegypti_ genes, first load relevant libraries and get the genes biomart (see my other tutorial for more info on the biomart):
__Note that the biomart name will change in the future so may need to be updated__

```{r cache = TRUE}
library(biomaRt)
aaegGenes<-useMart(biomart = "vb_gene_mart_1608", dataset = "aaegypti_eg_gene", host = "biomart.vectorbase.org")
```

## Get a list of all Aedes aegypti genes

```{r cache = T}
totalgenes<-getBM(
  attributes = c("ensembl_gene_id"),
  mart = aaegGenes)
```

## Get all the GO IDs and names associated with those genes

```{r cache = T}
goids<-getBM(
  # These are the attributes you want returned as columns in your data frame
  attributes = c("go_accession", "go_name_1006", "ensembl_gene_id"),
  # The filters is what attribute you are filtering on
  filters = "ensembl_gene_id", 
  # values is your vector of gene IDs
  values = totalgenes$ensembl_gene_id,
  # mart is your presaved BioMart database
  mart = aaegGenes)

```

Here you could also choose to filter based on GO evidence code by adding `"go_evidence_code"` as a filter and create a list from the following:  
`"IC"` - Inferred by curator  
`"IDA"` - Inferred by direct assay  
`"IEA"` - Inferred from electronic annotation  
`"IMP"` - Inferred from mutant phenotype  
`"ISS"` - Inferred from Sequence of structural similarity  

## Transform table from one to one to one GOID to many genes format

```{r cache = T}
# Replace all the spaces with underscores
goids$go_name_1006<-gsub(pattern = " ", replacement = "_", x = goids$go_name_1006)
# Get Unique list of GOIDs
GOAssoc<-goids[,1:2]
# Remove duplicates
GOAssoc<-GOAssoc[!duplicated(GOAssoc),]
# Remove blank
GOAssoc<-GOAssoc[GOAssoc$go_accession != "",]
```

Now we'll use a function to collect all Gene names associated to each GO term and store in

```{r, cache = T}
GOIDGeneLookup<-function(uniqueGOtable, fullAnnTable){
  GOTableLength<-length(uniqueGOtable[,1])
  # Create a vector to store results that is the same length as the GO term table
  geneids<-1:GOTableLength
  # Iterate over each goterm and store string of Gene IDs associated to that term
  for (i in 1:GOTableLength){
    GOID<-GOAssoc$go_accession[i]
    geneids[i]<-paste(goids$ensembl_gene_id[which(goids$go_accession == GOID)], collapse = ' ')
  }
  # Merge the ids to the rest of the table and return the new dataframe
  GowindaTable<-data.frame(uniqueGOtable, geneids)
  return(GowindaTable)
}

# Call the function on the data frames we have already created
GOWTable<-GOIDGeneLookup(GOAssoc, goids)
# Now we have our table in the right format for Gowinda
head(GOWTable[,1:2])
# Just need to save it as a file as follows:
write.table(x = GOWTable, file = "GowindaGeneSetFile.txt", quote = F, sep = "\t", col.names = F, row.names = F)

```

The outputted `GowindaGeneSetFile.txt` can now be passed to Gowinda with GO terms from the latest BioMart information.

Back to homepage: <https://mshadbolt.github.io/Ae_aegypti-toolset/>